---
title: "bootstrapping"
author: "Jesus Luevano"
date: "2023-11-16"
output: html_document
---

```{r}
library(tidyverse)
library(modelr)
library(p8105.datasets)
```


## Generate a relevant example

```{r}
n_samp = 250

sim_df_const = 
  tibble(
    x = rnorm(n_samp, 1, 1),
    error = rnorm(n_samp, 0, 1),
    y = 2 + 3 * x + error
  )

sim_df_nonconst = 
  sim_df_const %>%
  mutate(
    error = error * 0.75 * x, #x values close to 0 lower error, vice versa, so error based on x
    y = 2 + 3 * x + error
  )

sim_df_const %>%
  ggplot(aes(x = x, y = y)) + geom_point()

sim_df_nonconst %>%
  ggplot(aes(x = x, y = y)) + geom_point()
```

fit some linear models

```{r}
sim_df_const %>%
  lm(y ~ x, data = .) %>%
  broom::tidy()

sim_df_nonconst %>%
  lm(y ~ x, data = .) %>%
  broom::tidy()
```


## Draw and analyze a boostrap sample

Start w/ a lil function

```{r}
boot_sample = function(df){
  
  sample_frac(df, size = 1, replace = TRUE) #size =1 means draw same size as DF, so 100%
  
}
```

Let's see how this works

```{r}
sim_df_nonconst %>%
  boot_sample() %>%
  ggplot(aes(x = x, y = y)) + 
  geom_point(alpha = 0.5) + 
  stat_smooth(method = "lm")
```


```{r}
boot_straps =
  tibble(strap_number = 1:100) %>%
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(sim_df_nonconst)) #run over i, iterations of strap_number, for this anonymous function (i.e. just a normal command)
  )

boot_straps %>%
  pull(strap_sample) %>%
  nth(1) %>%
  arrange(x)
```

Now do the 'lm' fit.

```{r}
boot_results = 
  boot_straps %>%
  mutate(
    models = map(strap_sample, \(df) lm(y ~ x, data = df)),
    results = map(models, broom::tidy)
  ) %>%
  select(strap_number, results) %>%
  unnest(results)
```


Try to summarize these results -- get a bootstrap SE

```{r}
boot_results %>%
  group_by(term) %>%
  summarize(
    se = sd(estimate)
  )
```

```{r}
boot_results %>%
  filter(term == "x") %>%
  ggplot(aes(x = estimate)) +
  geom_density()
```

Can i Construct a CI

```{r}
boot_results %>%
  group_by(term) %>%
  summarize(
    ci_lower = quantile(estimate, 0.025), 
    ci_upper = quantile(estimate, 0.975)
  )
```


## Airbnb

```{r}
data("nyc_airbnb")

nyc_airbnb = nyc_airbnb %>%
  mutate(stars = review_scores_location /2) %>%
  rename(
    borough = neighbourhood_group
  ) %>%
  filter(borough != "Staten Island") %>%
  drop_na(price, stars) %>%
  select(price, stars, room_type, borough)


```

Let's fit a regression of 'price' on other variables and look at residuals

```{r}
airbnb_fite = 
  nyc_airbnb %>%
  lm(price ~ stars + room_type + borough, data = .)
```

residuals

```{r}
nyc_airbnb %>%
  modelr::add_residuals(airbnb_fite) %>%
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```

run a bootstrap on this whole thing to get estimates for the effect of 'stars' on 'price'

```{r}
manhattan_df = 
  nyc_airbnb %>%
  filter(borough == "Manhattan")

boot_results = 
  tibble(strap_number = 1:1000) %>%
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(manhattan_df)),
    models = map(strap_sample, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)
  ) %>%
  select(strap_number, results) %>%
  unnest(results)

boot_results %>%
  filter(term == "stars") %>%
  ggplot(aes(x = estimate)) +
  geom_density()
```

