---
title: "Cross Validation"
author: "Jesus Luevano"
date: "2023-11-14"
output: html_document
---

```{r}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

## Nonlinear data and CV

```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100,0,1),
    y = 1 - 10 * (x - 0.3)^2 + rnorm(100,0,0.3)
  )

nonlin_df %>%
  ggplot(aes(x = x, y = y)) + 
  geom_point()
```

Do the train/test split.

```{r}
train_df = sample_n(nonlin_df, 80)
test_df  = anti_join(nonlin_df, train_df, by = "id")
```

```{r}
train_df %>%
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_point(data = test_df, color = "red")
```

```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = mgcv::gam(y ~ s(x), data = train_df)
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

Quick visualization

```{r}
train_df %>%
  modelr::add_predictions(linear_mod) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_path(aes(y=pred))

train_df %>%
  modelr::add_predictions(smooth_mod) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y=pred))

train_df %>%
  modelr::add_predictions(wiggly_mod) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y=pred))
```

To compare them to see which best, as not nested can't compare directly or use F-test or AIC/BIC. So need to try something else


But RMSE on training data can be isleading
```{r}
rmse(linear_mod, train_df) #but if don't set seed, these vary
rmse(smooth_mod, train_df)
rmse(wiggly_mod, train_df)
```

RMSE on testing data gives sense of out-of-sample prediciton accuracy!

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)

```


## Use mdelr for CV

```{r}
cv_df =
  nonlin_df %>%
  crossv_mc(n = 100) %>% # saves as list of dataframes
  mutate(
    train = map(train, as_tibble),
    test = map(test,as_tibble)
  )
```

```{r}
cv_df %>% pull(train) %>% nth(3) %>% as_tibble() #re-sampling instead
```

Apply each model to all training datasets, and evaluate on all testing datasets.

```{r}
cv_results = 
  cv_df %>%
  mutate(
    linear_fit = map(train, \(df) lm(y ~ x, data = df)),
    smooth_fit = map(train, \(df) mgcv::gam(y ~ s(x), data = df)),
    wiggly_fit = map(train, \(df) mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = df))
  ) %>%
  mutate(
    rmse_lienar = map2_dbl(linear_fit, test, \(mod, df) rmse(mod,df)),
    rmse_smooth = map2_dbl(smooth_fit, test, \(mod, df) rmse(mod, df)),
    rmse_wiggly = map2_dbl(wiggly_fit, test, \(mod, df) rmse(mod, df))
  )
```

```{r}
cv_results %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model_type",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>%
  group_by(model_type) %>%
  summarize(m_rmse = mean(rmse))
```

